---
title: "Gas Price Sub Markets"
author: "Kai Lukowiak"
date: '2018-12-10'
output: ioslides_presentation
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

```{r  warning=FALSE, message=FALSE}
source('function.R')
source('data_manipulator.R')
source('knn.R')
```

## Thesis

*To what extent are submarkets in gasoline prices segmented accross distinct geographic teritories?*

## Problem Overview

Gasoline markets are very local. While there is correlation in price over time between cities this is almost entirely due to crude oil and refining costs. NYC is correlated with Calgary to some extent, but not when you look at margins.

These costs, including transport and called the "rack" in Canada. The (retail price) - rack is the margin. Because the rack to retail (R2R) is independant of the rack, margins are uncorrelated between cities.

## Gas Markets
- Gas prices are highly competitive
- The price payed for gasoline only contains a bit of profit for the station
- Margins of 0 to 10% are the norm. 
- There is significant fluctiation throught the day and year

## Submarkets
Markets become much less correlated when you compare margins than when you compare prices.

```{r   warning=FALSE, message=FALSE}
library(ggplot2)
library(gridExtra)

p1 <- dt[,
         .(margin = mean(margin, na.rm = T)),
         by = .(date(time), area_name)] %>% 
  ggplot(aes(date, margin, color = area_name)) +
  geom_line() + 
  theme_classic() +
  theme(legend.position="bottom")
p2 <- dt[,
         .(price = mean(price, na.rm = T)),
         by = .(date(time), area_name)] %>% 
  ggplot(aes(date, price, color = area_name)) +
  geom_line()+ 
  theme_classic()+
  theme(legend.position="bottom")

grid.arrange(p1, p2, nrow = 1)
```


## Correlation
**Prices**
```{r warning=F, message=F}
library(Hmisc)
temp = dt[, .(mean_price = mean(price, na.rm = T), 
              mean_margin = mean(margin, na.rm = T)), 
          by = .(area_name, date(time))]
price_cor <- dcast(data = temp,
      formula = date~area_name,fun.aggregate = mean, value.var = 'mean_price')

price_cor <- price_cor[complete.cases(price_cor)]

margin_cor <- dcast(data = temp,
      formula = date~area_name,fun.aggregate = mean, value.var = 'mean_margin')
margin_cor = margin_cor[complete.cases(margin_cor)]

price_cor <- rcorr(as.matrix(price_cor[, -'date']))
round(price_cor$r, 2)
margin_cor <- rcorr(as.matrix(margin_cor[, -'date']))

```


## Correlation
**Margins**
```{r  warning=FALSE, message=FALSE}
round(margin_cor$r, 2)
```

so while there is some correlation, it's not nearly as distinct as with prices.


## Existing Literature

Much work has been done analyzing gasoline markets because they are highly visible and people spend a lot of money on gas.

Existing literature focused on small sample sizes (although considerably more accurate than mine) and spatial regression.

## My Approach
### Python
- Scrape GasBuddy.com
  - Python/AWS/Selenium/Lots of errors
- Pull location info from google maps API
  - I had grandiouse ideas to scrape weather and traffic but due to budget limitations, I did not
  - Python
- Manipulate data using Pandas Resample

## My Approach
### R
I used data.table and TSclust to cluster time series data using correlation and eucludian distance. I then picked the number of clusters and used this as 'arbitrary' labes.

I then split this into train and test sets and performed KNN classification on this using latitude and longitude.

## Graphs
```{r  warning=FALSE, message=FALSE}
theme_set(theme_classic())
p1 <- calg_lab %>% 
  ggplot(aes(lat, lng, color = calg_raw_cor_2))+
  geom_point() + theme(legend.position="none") + 
  ggtitle("Raw Corelated 2 Calgary")

p2 <- calg_lab %>% 
  ggplot(aes(lat, lng, color = calg_adj_cor_2))+
  geom_point() + theme(legend.position="none") +
  ggtitle('Adjusted Correlated 2 Calgary')

p3 <- calg_lab %>% 
  ggplot(aes(lat, lng, color = calg_raw_eucl_2))+
  geom_point() + theme(legend.position="none") +
  ggtitle("Raw Eucl 2 Calgary")

p4 <- calg_lab %>% 
  ggplot(aes(lat, lng, color = calg_adj_eucl_2))+
  geom_point() + theme(legend.position="none") +
  ggtitle("Adj Eucl 2 Calgary")
pdl <- grid.arrange(p1, p2, p3, p4, nrow = 2)

```

## Calgary 5 Sites
```{r  warning=FALSE, message=FALSE}
theme_set(theme_classic())
p1 <- calg_lab %>% 
  ggplot(aes(lat, lng, color = calg_raw_cor_5))+
  geom_point() + theme(legend.position="none") + 
  ggtitle("Raw Corelated 5 Calgary")

p2 <- calg_lab %>% 
  ggplot(aes(lat, lng, color = calg_adj_cor_5))+
  geom_point() + theme(legend.position="none") +
  ggtitle('Adjusted Correlated 5 Calgary')

p3 <- calg_lab %>% 
  ggplot(aes(lat, lng, color = calg_raw_eucl_5))+
  geom_point() + theme(legend.position="none") +
  ggtitle("Raw Eucl 5 Calgary")

p4 <- calg_lab %>% 
  ggplot(aes(lat, lng, color = calg_adj_eucl_5))+
  geom_point() + theme(legend.position="none") +
  ggtitle("Adj Eucl 5 Calgary")
grid.arrange(p1, p2, p3, p4, nrow = 2)

```

## Western Canada Combined
```{r  warning=FALSE, message=FALSE}
p1 <- calg_lab %>% 
  ggplot(aes(lat, lng, color = calg_raw_eucl_2))+
  geom_point() + theme(legend.position="none") + ggtitle("Raw Euclid 2 Calgary")

p2 <- calg_lab %>% 
  ggplot(aes(lat, lng, color = calg_adj_eucl_2)) +
  geom_point() + theme(legend.position="none") +
  ggtitle('Adjusted Euclid 2 Calgary')

p3 <- calg_lab %>% 
  ggplot(aes(lat, lng, color = calg_adj_eucl_5)) +
  geom_point() + theme(legend.position="none") +
  ggtitle("Adjusted Euclid 5 Calgary")

p4 <- all_lab[area_name == 'calgary'] %>% 
  ggplot(aes(lat, lng, color = adj_eucl_6)) +
  geom_point() + theme(legend.position="none") +
  ggtitle("All Adjusted Euclid 6")
grid.arrange(p1, p2, p3, p4, nrow = 2)

```


## 
```{r warning=FALSE, message=FALSE}
p1 <- calg_lab %>% 
  ggplot(aes(lat, lng, color = calg_raw_cor_2))+
  geom_point() + theme(legend.position="none") + ggtitle("Raw Cor 2 Calgary")

p2 <- calg_lab %>% 
  ggplot(aes(lat, lng, color = calg_adj_cor_2))+
  geom_point() + theme(legend.position="none") +
  ggtitle('Adjusted Correlated 2 Calgary')

p3 <- calg_lab %>% 
  ggplot(aes(lat, lng, color = calg_adj_cor_5))+
  geom_point() + theme(legend.position="none") +
  ggtitle("Adjusted Cor 5")

p4 <- all_lab[area_name == 'calgary'] %>% 
  ggplot(aes(lat, lng, color = adj_cor_12))+
  geom_point() + theme(legend.position="none") +
  ggtitle("All Adjusted Correlated 12")
grid.arrange(p1, p2, p3, p4, nrow = 2)

```

## Best Clustering

```{r  warning=FALSE, message=FALSE}
p1 <- all_lab[area_name == 'calgary'] %>% 
  ggplot(aes(lat, lng, color = raw_cor_6)) +
  geom_point() + theme(legend.position="none") + 
  ggtitle("All Raw Cor 6")

p2 <- all_lab[area_name == 'calgary'] %>% 
  ggplot(aes(lat, lng, color = raw_eucl_6))+
  geom_point() + theme(legend.position="none") +
  ggtitle('All Raw Eucl 6')

p3 <- all_lab[area_name == 'calgary'] %>% 
  ggplot(aes(lat, lng, color = adj_cor_6))+
  geom_point() + theme(legend.position="none") +
  ggtitle("All Adjusted Euclid 6")

p4 <- all_lab[area_name == 'calgary'] %>% 
  ggplot(aes(lat, lng, color = adj_eucl_6))+
  geom_point() + theme(legend.position="none") +
  ggtitle("All Adjusted Eucl 6")
grid.arrange(p1, p2, p3, p4, nrow = 2)
```

## Comparison to Vancouver

```{r warning=FALSE, message=FALSE}
p1 <- all_lab[area_name == 'vancouver'] %>% 
  ggplot(aes(lat, lng, raw, color = raw_eucl_12)) +
  geom_point() +  theme(legend.position="none") +
  ggtitle("All Cor 6 (van)")
p2 <- van_lab %>% 
  ggplot(aes(lat, lng, color = van_adj_cor_5)) +
  geom_point() +  theme(legend.position="none") +
  ggtitle('Vancouver Adj Cor 2')

grid.arrange(p1, p2, nrow = 1)
```

## Selection
- I will test more exhaustively in the final project
- Current analysis focuses on All Raw Euclidian 6.

*How does this perform?*

```{r  warning=FALSE, message=FALSE}
mod1 <- knn_maker(all_lab[, .(raw_eucl_6, lat, lng)])
mod1$accuracy_table
```


## Confusion Matrix

```{r}
confusionMatrix(mod1$accuracy_table)
```



## Next Best Alternative
Taking the most common lable by city:
```{r}
mode_clust <- all_lab %>% 
  dplyr::as_data_frame() %>% 
  dplyr::group_by(area_name) %>% 
  dplyr::count(raw_eucl_6, area_name) %>%
  dplyr::slice(which.max(n)) 
mode_clust <- data.table(mode_clust)

next_best <- merge(all_lab, mode_clust, by = 'area_name')
# next_best <- next_best[raw_eucl_6.x != 'Cluster_4' & raw_eucl_6.x != "Cluster_6"]
next_best <- table(next_best[,.(raw_eucl_6.x, raw_eucl_6.y)])
next_best
acc <- (107+67+165+32) / sum(colSums(next_best))
print(paste("The accuracy is:", acc))
# confusionMatrix(next_best)
# all_lab[, .(count = uniqueN(area_name)), by = .(raw_eucl_6)]
```

## Results
**Not Great** 

These results are not incredibly promissing although they do fall within of the 95% CI for the accuracy measure. 

## Individal Cities

Taking the most promissing result for Calgary, the adjusted correlation 2
```{r}
mod1 <- knn_maker(calg_lab[, .(calg_adj_cor_2, lat, lng)])
# print("The tuning K for best CV accuracy is:")
# mod1$fit$bestTune
# print("The Accuracy Table is:")
mod1$accuracy_table
```


## Analysis
If we chose the most common lable in calgary we get:

```{r}
calg_lab[, .(count_calg = .N), by = calg_adj_cor_2]
```

This would only give us an accuracy of `r 84/(74+82)`.

This suggests that aggrigating over entire markets only captures the market, not the sub markets.


## Increase K
If we increase K to 5:
```{r}
mod1 <- knn_maker(calg_lab[, .(calg_adj_cor_5, lat, lng)])
# print("The tuning K for best CV accuracy is:")
# mod1$fit$bestTune
print("The Accuracy Table is:")
mod1$accuracy_table
```

## Confusion Matrix
For K=5 for Calgary
```{r}
confusionMatrix(mod1$accuracy_table)

```


## Analysis

If we chose the most common lable in calgary we get:

```{r}
calg_lab[, .(count_calg = .N), by = calg_adj_cor_5]
```


This would give an accuracy of `r 71/ calg_lab[, .N]` which is greater than the KNN accuracy. 
This leads me to believe that sub markets within cities do not exist.

## Issues
There are several issues that I faced that might hide a different result.

- The data collection was spoty due to AWS
- GasBuddy does not have very granular pricing
- Since no company to my knowlege tries to set prices at the neighborhood level there could still be potential gains from trying too.

## Results
While the cluster analysis worked well defining different cities, it was a poor predictor of within city markets.

**Explanation**

Nobody travels accross town for gas (Costco is an exception), but it is only marginally more work to move out of a neighborhood, especially since people commute great distances. Thus, submarkets merge accross cities.

## Further Work

See hierarchies